### 2021-04-08 ch03_esmvaltool/esmvaltool.tex
Standardized use of indefinite articles with acronyms
Spell- and grammer-checked chapter 1

To address the \namecrefs{enum:01:question_1} of this thesis posed in
\cref{sec:01:key_science_questions}, a reliable and efficient tool to read,
process, and evaluate climate model output and observational data is necessary.
A valuable software that fits these criteria is the \ac{ESMValTool}. The
\ac{ESMValTool} is an open-source community diagnostics and performance metrics
tool for the routine evaluation of \ac{ESM} output, which notably facilitates
the analysis of \ac{CMIP} models (\url{https://www.esmvaltool.org}). For this
reason, all analyses presented in
\cref{ch:04:papers_ecs_tcr_assessment,ch:05:paper_ecs,ch:06:paper_gpp} of this
thesis have been implemented into the \ac{ESMValTool}. Apart from that, further
substantial changes and additions to the code base of the \ac{ESMValTool} have
been contributed that improve the routine evaluation of climate models, which
is beneficial for the entire scientific community. This led to co-authorship in
the scientific documentation of the \ac{ESMValTool}, which is published in four
peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
After a brief overview of the \ac{ESMValTool}'s structure
(\cref{sec:03:esmvaltool}), this \namecref{ch:03:esmvaltool} presents all these
contributions that are not documented in other
\namecrefs{ch:04:papers_ecs_tcr_assessment} of this thesis and puts them into
context
(\cref{sec:03:contributions_to_esmvalcore,sec:03:contributions_to_esmvaltool}).

@@

Due to an exhaustive overhaul of the \ac{ESMValTool} version 2, older recipes
from version 1 \autocite{Eyring2016a} cannot be used directly in the latest
release. Porting recipes from the old version to the current version requires
substantial changes on the recipes themselves, but also on the corresponding
diagnostic scripts. As of \TheMonth{}, the author of this thesis has ported two
recipes originally written by other authors to the new version of
\ac{ESMValTool}: \emph{recipe\_anav13jclim.yml} and
\emph{recipe\_wenzel16jclim.yml}. \emph{recipe\_anav13jclim.yml} reproduces the
analysis of \textcite{Anav2013}, who evaluate carbon cycle--related variables
for \acp{ESM} of the \acs{CMIP}5 ensemble. This includes climatologies, trends,
and variabilities of important quantities of the terrestrial and oceanic carbon
cycle like carbon fluxes, carbon reservoir sizes, and vegetation distributions.
\emph{recipe\_wenzel16jclim.yml} includes the \ac{MDER} analysis performed by
\textcite{Wenzel2016a} to constrain future changes in the position of the
austral jet stream. Due to its flexible code, the corresponding diagnostic
scripts can be used to apply the \ac{MDER} method to arbitrary target variables
and predictors. Both recipes include analyses that contributed ideas and
methods for the study presented in \cref{ch:06:paper_gpp}, in which the
\ac{MDER} method is expanded with a \ac{ML} approach and applied to a carbon
cycle--related target variable (\ac{GPP}).

@@

An important class of recipes in the \ac{ESMValTool} covers international
climate assessments. One such example is \emph{recipe\_flato13ipcc.yml}, which
reproduces large parts of the climate model evaluation performed in chapter 9
of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
recipe, the two-panel figure 9.42 has been added, which shows the relationship
between the historical and pre-industrial \ac{GSAT} and \ac{ECS}
(\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS}
(\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble.
\Cref{fig:03:flato13:a} illustrates that there is no clear connection between
\ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high
temperatures in the historical or pre-industrial period do not necessarily
exhibit a high equilibrium warming in the future or vice versa. On the other
hand, \ac{TCR} and \ac{ECS} are in fact well correlated within the \acs{CMIP}5
ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this
figure only shows a linear regression fit between the two variables and not the
expected reciprocal relation. The output of \emph{recipe\_flato13ipcc.yml}
provided useful input for the analysis of the climate sensitivity in
\acs{CMIP}6 performed in \textcite{Meehl2020}, which is also presented in
\cref{ch:04:papers_ecs_tcr_assessment}.

>>>>

To address the \namecrefs{enum:01:question_1} of this thesis posed in
\cref{sec:01:key_science_questions}, a reliable and efficient tool to read,
process, and evaluate climate model output and observational data is necessary.
A valuable software package that fits these criteria is the \ac{ESMValTool}.
The \ac{ESMValTool} is an open-source community diagnostics and performance
metrics tool for the routine evaluation of \ac{ESM} output, which notably
facilitates the analysis of \ac{CMIP} models
(\url{https://www.esmvaltool.org}). For this reason, all analyses presented in
\cref{ch:04:papers_ecs_tcr_assessment,ch:05:paper_ecs,ch:06:paper_gpp} of this
thesis have been implemented into the \ac{ESMValTool}. Apart from that, further
substantial changes and additions to the code base of the \ac{ESMValTool} have
been contributed that improve the routine evaluation of climate models, which
is beneficial for the entire scientific community. This led to co-authorship in
the scientific documentation of the \ac{ESMValTool}, which is published in four
peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
After a brief overview of the \ac{ESMValTool}'s structure
(\cref{sec:03:esmvaltool}), this \namecref{ch:03:esmvaltool} presents all these
contributions that are not documented in other
\namecrefs{ch:04:papers_ecs_tcr_assessment} of this thesis and puts them into
context
(\cref{sec:03:contributions_to_esmvalcore,sec:03:contributions_to_esmvaltool}).

@@

Due to an exhaustive overhaul of the \ac{ESMValTool} version 2, older recipes
from version 1 \autocite{Eyring2016a} cannot be used directly in the latest
release. Porting recipes from the old version to the current version requires
substantial changes on the recipes themselves, but also on the corresponding
diagnostic scripts. As of \TheMonth{}, the author of this thesis has ported two
recipes originally written by other authors to the new version of
\ac{ESMValTool}: \emph{recipe\_anav13jclim.yml} and
\emph{recipe\_wenzel16jclim.yml}. \emph{recipe\_anav13jclim.yml} reproduces the
analysis of \textcite{Anav2013}, who evaluate carbon cycle--related variables
for \acp{ESM} of the \acs{CMIP}5 ensemble. This includes climatologies, trends,
and variabilities of important quantities of the terrestrial and oceanic carbon
cycle like carbon fluxes, carbon reservoir sizes, and vegetation distributions.
\emph{recipe\_wenzel16jclim.yml} includes the \ac{MDER} analysis performed by
\textcite{Wenzel2016a} to constrain future changes in the position of the
austral jet stream. Due to its flexible code, the corresponding diagnostic
scripts can be used to apply the \ac{MDER} method to arbitrary target variables
and predictors. Both recipes include analyses that contributed ideas and
methods for the study presented in \cref{ch:06:paper_gpp}, in which the
\ac{MDER} method is expanded with \iac{ML} approach and applied to a carbon
cycle--related target variable (\ac{GPP}).

@@

An important class of recipes in the \ac{ESMValTool} covers international
climate assessments. One such example is \emph{recipe\_flato13ipcc.yml}, which
reproduces large parts of the climate model evaluation performed in chapter 9
of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
recipe, the two-panel figure 9.42 has been added, which shows the relationship
between the historical and pre-industrial \ac{GSAT} and \ac{ECS}
(\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS}
(\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble.
\Cref{fig:03:flato13:a} illustrates that there is no clear connection between
\ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high
temperatures in the historical or pre-industrial period do not necessarily
exhibit a high equilibrium warming in the future or vice versa. On the other
hand, \ac{TCR} and \ac{ECS} are in fact well correlated within the \acs{CMIP}5
ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this
figure only shows a linear regression fit between the two variables and not the
expected reciprocal relation. The output of \emph{recipe\_flato13ipcc.yml}
provided useful input for the analysis of climate sensitivity in \acs{CMIP}6
performed in \textcite{Meehl2020}, which is also presented in
\cref{ch:04:papers_ecs_tcr_assessment}.

### 2021-04-08 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Spell- and grammer-checked chapter 1

To illustrate the results of the previous
\namecref{sec:04:evaluation_ecs_and_tcr} and put them into historical context,
\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
and the different \acp{AR} of the \ac{IPCC} in combination with the
corresponding modeled ranges from the different \ac{CMIP} generations. Since
the Charney report in 1979, the assessed range of \ac{ECS} of
$\rangeunit{1.5}{4.5}{K}$ has almost remained unchanged during the last 40
years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
  Stocker2013} with the exception of \acs{AR}4, where the lower bound was
temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
\acs{AR}4, and \acs{AR}5, the corresponding climate model generations from
\acs{CMIP}1, \acs{CMIP}3, and \acs{CMIP}5 more or less agree with this assessed
range \autocite{Kattenberg1996, Randall2007, Flato2013}. In contrast to that,
some of the early climate models used in \acs{AR}1 and some \acs{CMIP}2 models
used in \acs{AR}3 exhibit values of \acs{ECS} well above $4.5 \unit{K}$,
resulting in an upper model range of about $0.6 \unit{K}$ higher than the
assessed range \autocite{Mitchell1990, Cubasch2001}. However, these deviations
from the assessed range are small compared to the ones present in the
\acs{CMIP}6 ensemble, which have become evident over the course of 2020. As
shown in \cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper
bound are more extreme than in any previous climate model generation, resulting
in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$ (as of
\TheMonth{}). This increase in range is particularly relevant for the upper
bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that exceed
this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third (14
models) have \iac{ECS} above $4.5 \unit{K}$. In addition to the analysis
presented here, several modeling centers independently confirmed high \ac{ECS}
values in their models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019,
  Wyser2020}. A list of possible reasons for the increased climate sensitivity
in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes
in the shortwave cloud feedback over the Southern Ocean is given in
\cref{sec:04:possible_reasons_high_ecs_cmip6}.

@@

Because of the massive political and societal relevance of \ac{ECS}, its
apparent increase in the \acs{CMIP}6 ensemble is currently one of the most
important questions for the entire climate modeling community. For this reason,
this \namecref{sec:04:possible_reasons_high_ecs_cmip6} discusses possible
reasons for the increased climate sensitivity in \acs{CMIP}6.

@@

Another possible reason for the high climate sensitivity in the \acs{CMIP}6
ensemble is related to cloud feedbacks \autocite{Bock2020}. As already
discussed in \cref{subsec:02:climate_feedbacks}, uncertainties in cloud
feedbacks are a major source of uncertainty of \ac{ECS} in modern \acp{ESM}
\autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
to cloud feedbacks immediately impact the climate models' climate sensitivity.
The particular feedback mechanism that is thought to be relevant for the high
\ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5).
This so-called \emph{cloud phase change feedback} is illustrated in
\cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate warms,
the predominantly ice clouds over the Southern Ocean in these climate models
become liquid clouds \autocite{McCoy2015}. As a result, the cloud will get more
reflective to the incoming solar radiation since a cloud consisting of smaller
liquid droplets reflects more sunlight than a cloud consisting of larger ice
crystals (for a fixed water content), which has a cooling effect on the climate
system. In addition, predominantly liquid clouds tend to precipitate less than
mixed-phase clouds (composed of liquid water and ice). This leads to a higher
liquid water content of the cloud and a further amplification of the cooling
effect. Overall, this forms a strong negative shortwave cloud feedback over the
Southern Ocean in climate models from previous \ac{CMIP} generations.

>>>>

To illustrate the results of the previous
\namecref{sec:04:evaluation_ecs_and_tcr} and put them into historical context,
\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
and the different \acp{AR} of the \ac{IPCC} in combination with the
corresponding modeled ranges from the different \ac{CMIP} generations. Since
the Charney report in 1979, the assessed range of \ac{ECS} of
$\rangeunit{1.5}{4.5}{K}$ has almost remained unchanged during the last 40
years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
  Stocker2013} with the exception of \acs{AR}4, where the lower bound was
temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
\acs{AR}4, and \acs{AR}5, the corresponding climate model generations from
\acs{CMIP}1, \acs{CMIP}3, and \acs{CMIP}5 more or less agree with this assessed
range \autocite{Kattenberg1996, Randall2007, Flato2013}. In contrast to that,
some of the early climate models used in \acs{AR}1 and some \acs{CMIP}2 models
used in \acs{AR}3 exhibit values of \acs{ECS} well above $4.5 \unit{K}$,
resulting in an upper model range of about $0.6 \unit{K}$ higher than the
assessed range \autocite{Mitchell1990, Cubasch2001}. However, these deviations
from the assessed range are small compared to the ones present in the
\acs{CMIP}6 ensemble, which have become evident over the course of 2020. As
shown in \cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper
bound are more extreme than in any previous climate model generation, resulting
in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$ (as of
\TheMonth{}). This increase in range is particularly relevant for the upper
bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that exceed
this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third (14
models) have \iac{ECS} above $4.5 \unit{K}$. In addition to the analysis
presented here, several modeling centers independently confirmed high \ac{ECS}
values in their models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019,
  Wyser2020}. A list of possible reasons for increased climate sensitivity in
the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
the shortwave cloud feedback over the Southern Ocean is given in
\cref{sec:04:possible_reasons_high_ecs_cmip6}.

@@

Because of the massive political and societal relevance of \ac{ECS}, its
apparent increase in the \acs{CMIP}6 ensemble is currently one of the most
important questions for the entire climate modeling community. For this reason,
this \namecref{sec:04:possible_reasons_high_ecs_cmip6} discusses possible
reasons for increased climate sensitivity in \acs{CMIP}6.

@@

Another possible reason for high climate sensitivity in the \acs{CMIP}6
ensemble is related to cloud feedbacks \autocite{Bock2020}. As already
discussed in \cref{subsec:02:climate_feedbacks}, uncertainties in cloud
feedbacks are a major source of uncertainty of \ac{ECS} in modern \acp{ESM}
\autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
to cloud feedbacks immediately impact climate sensitivity of the climate
models. The particular feedback mechanism that is thought to be relevant for
the high \ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the
Southern Ocean that are present in earlier \ac{CMIP} generations (\eg{}
\acs{CMIP}5). This so-called \emph{cloud phase change feedback} is illustrated
in \cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate
warms, the predominantly ice clouds over the Southern Ocean in these climate
models become liquid clouds \autocite{McCoy2015}. As a result, the cloud will
get more reflective to the incoming solar radiation since a cloud consisting of
smaller liquid droplets reflects more sunlight than a cloud consisting of
larger ice crystals (for a fixed water content), which has a cooling effect on
the climate system. In addition, predominantly liquid clouds tend to
precipitate less than mixed-phase clouds (composed of liquid water and ice).
This leads to a higher liquid water content of the cloud and a further
amplification of the cooling effect. Overall, this forms a strong negative
shortwave cloud feedback over the Southern Ocean in climate models from
previous \ac{CMIP} generations.

### 2021-04-08 ch07_conclusion/conclusion.tex
Spell- and grammer-checked chapter 1

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS}, a
\acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is higher
on the upper and lower bound than any model range from previous \ac{CMIP}
generations. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of \ac{ECS}
is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus $3.23
\unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
\acs{CMIP}6 model range is considerably larger than in the assessed range of
$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
\ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
\unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with many models
showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model range of
\acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the \acs{CMIP}5
range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased climate
sensitivity in some \acs{CMIP}6 models is the addition of prognostic aerosol
schemes that include aerosol-cloud interactions, which might result in overly
large negative radiative forcing. In this case, a stronger response to \ac{GHG}
forcing is required to correctly reproduce the observed historical temperature
trend, resulting in a higher climate sensitivity in the corresponding models. A
further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in the
microphysical representation of mixed-phase clouds in some \acs{CMIP}6 models.
This change was implemented to improve the representation of the fractioning of
cloud ice and cloud liquid in those models, \eg{} by allowing for (more)
supercooled cloud liquid water. In addition, the \acs{CMIP}6 \ac{MMM} shows an
improved simulation of the shortwave \ac{CRE} when compared to satellite
observations. The change in the representation of mixed-phase clouds in some
\acs{CMIP}6 models reduces the strong negative shortwave cloud feedback over
the Southern Ocean that is present in predecessor versions of these models and
that results from a cloud phase change from ice clouds in the present-day to
liquid clouds in the future. In the affected \acs{CMIP}6 models, this cloud
phase change due to warming is reduced since these models simulate less cloud
ice over the present-day Southern Ocean than their predecessor versions with no
supercooled cloud liquid.

@@

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \Ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, a
published emergent constraint by \textcite{Wenzel2016} is used to constrain the
global mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
unconstrained \acs{CMIP}5 model range of $\rangeunit{156}{247}{\GtCyr{}}$. This
first step corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2},
which is primarily driven by the \ac{CO2} fertilization. In the second step, a
\acs{ML}-based climate model weighting approach is used to further constrain
the gridded \ac{GPP} based on present-day predictors that are relevant for the
simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach is mathematically
similar to the \ac{MDER} approach \autocite{Karpechko2013, Senftleben2020,
  Wenzel2016a}, but additionally considers multi-dimensional (gridded) target
variables and non-linear relationships between the predictors and the target
variables. A relationship between process-oriented predictors and future
projections of \ac{GPP} is established and then utilized to project today's
observed conditions into the future. The prediction phase of the new method can
be interpreted as an implicit performance weighting. However, due to the
complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
to extract specific values for the individual weights. Two target variables are
considered: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave-one-model-out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
and \ac{LAI} are the most important features. A corresponding feature map shows
that \ac{T} dominates in the tropics and \ac{LAI} dominates in the northern
high latitudes, which might be attributed to an additional greening trend in
the high latitudes that enhances the \ac{GPP} increase in this region compared
to regions closer to the equator through an extension of the growing season.
This study directly addresses \cref{enum:01:question_3}
(\enquote{\KeyScienceQuestionThree{}}), which can be confidently answered with
\enquote{yes} based on the results found.

>>>>

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS}, a
\acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is higher
on the upper and lower bound than any model range from previous \ac{CMIP}
generations. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of \ac{ECS}
is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus $3.23
\unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
\acs{CMIP}6 model range is considerably larger than in the assessed range of
$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
\ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
\unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with many models
showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model range of
\acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the \acs{CMIP}5
range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. One possible reason for increased climate
sensitivity in some \acs{CMIP}6 models is the addition of prognostic aerosol
schemes that include aerosol-cloud interactions, which might result in overly
large negative radiative forcing. In this case, a stronger response to \ac{GHG}
forcing is required to correctly reproduce the observed historical temperature
trend, resulting in a higher climate sensitivity in the corresponding models. A
further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in the
microphysical representation of mixed-phase clouds in some \acs{CMIP}6 models.
This change was implemented to improve the representation of the fractioning of
cloud ice and cloud liquid in those models, \eg{} by allowing for (more)
supercooled cloud liquid water. In addition, the \acs{CMIP}6 \ac{MMM} shows an
improved simulation of the shortwave \ac{CRE} when compared to satellite
observations. The change in the representation of mixed-phase clouds in some
\acs{CMIP}6 models reduces the strong negative shortwave cloud feedback over
the Southern Ocean that is present in predecessor versions of these models and
that results from a cloud phase change from ice clouds in the present-day to
liquid clouds in the future. In the affected \acs{CMIP}6 models, this cloud
phase change due to warming is reduced since these models simulate less cloud
ice over the present-day Southern Ocean than their predecessor versions with no
supercooled cloud liquid.

@@

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points to train the \ac{ML}
algorithm, scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is not
an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \Ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, a
published emergent constraint by \textcite{Wenzel2016} is used to constrain the
global mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
unconstrained \acs{CMIP}5 model range of $\rangeunit{156}{247}{\GtCyr{}}$. This
first step corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2},
which is primarily driven by the \ac{CO2} fertilization. In the second step, an
\acs{ML}-based climate model weighting approach is used to further constrain
the gridded \ac{GPP} based on present-day predictors that are relevant for the
simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach is mathematically
similar to the \ac{MDER} approach \autocite{Karpechko2013, Senftleben2020,
  Wenzel2016a}, but additionally considers multi-dimensional (gridded) target
variables and non-linear relationships between the predictors and the target
variables. A relationship between process-oriented predictors and future
projections of \ac{GPP} is established and then utilized to project today's
observed conditions into the future. The prediction phase of the new method can
be interpreted as an implicit performance weighting. However, due to the
complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
to extract specific values for the individual weights. Two target variables are
considered: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave-one-model-out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
and \ac{LAI} are the most important features. A corresponding feature map shows
that \ac{T} dominates in the tropics and \ac{LAI} dominates in the northern
high latitudes, which might be attributed to an additional greening trend in
the high latitudes that enhances the \ac{GPP} increase in this region compared
to regions closer to the equator through an extension of the growing season.
This study directly addresses \cref{enum:01:question_3}
(\enquote{\KeyScienceQuestionThree{}}), which can be confidently answered with
\enquote{yes} based on the results found.
